The role of lithic bipolar technology in Western Iberia’s Upper Paleolithic: the case of Vale Boi (southern Portugal)

### 2018-12-03 analysis/paper/paper.Rmd
morpho plot
inline statistics change

@@

As in other classes of stone tools, scaled pieces' metric attributes are impacted by both the initial blank size as well as by the intensity of their use. Specifically, because some scaled pieces are used in multiple axes, direct comparisons for length and width of the typological axes cannot be straightforwardly performed. Since the majority of the analyzed pieces presented a rectangular outline, we use the area of a rectangle (i.e. Length x Width) as an approximation for the overall dimensions of the artifacts. Area calculations revealed a maximum of `r round(max(techno_data$Area, na.rm = TRUE), digits = 2)` mm and a minimum of `r round(min(techno_data$Area, na.rm = TRUE), digits = 2)` mm. For thickness the maximum is `r round(max(techno_data$Thickness, na.rm = TRUE), digits = 2)` mm and the minimum is `r round(min(techno_data$Thickness, na.rm = TRUE), digits = 2)` mm. In general, mean values tend to be similar between raw materials, with some differences occurring within the Proto-Solutrean and Magdalenian assemblages, most certainly as a result of the small samples analysed for each of these periods (Figure \@ref(fig:meanplots)). Across techno-complexes, however, no significant statistical differences were detected when using an ANOVA test for both Area (*F* (`r aov_area$df[1]`, `r aov_area$df[2]`) = `r aov_area$statistic[1]`, *p* = `r aov_area$p.value[1]`, *d* = `r cohens_area$cohens.f`), and Thickness (*F* (`r aov_thickness$df[1]`, `r aov_thickness$df[2]`) = `r aov_thickness$statistic[1]`, *p* = `r aov_thickness$p.value[1]`, *d* = `r cohens_thickness$cohens.f`).

@@

In three (Scar Arrangement, Scar Extension and Edge Delineation) of the six represented variables, the category 'Other' is one of the most frequent (above 35%) for both quartz and chert, revealing a very high variability for the combination of attributes within each variable. A chi-square test with modified alpha level (Bonferroni correction) to adjust for multiple testing and reduce type-I error, shows significant differences in the quartz and chert Scar Arrangement categories (*X^2^* [`r chi_arrangement$chi_test$parameter`, *N* = `r sum(chi_arrangement$chi_test$observed)`] = `r  round(chi_arrangement$chi_test$statistic, 2)`, *p* = `r round(chi_arrangement$chi_test$p.value, 3)`, *phi* = `r round(chi_arrangement$chi_size_effect$phi, 2)`), and in Scar Faciality categories (*X^2^* [`r chi_faciality$chi_test$parameter`, *N* = `r sum(chi_faciality$chi_test$observed)`] = `r  round(chi_faciality$chi_test$statistic,2)`, *p* = `r round(chi_faciality$chi_test$p.value, 3)`, *phi* = `r round(chi_faciality$chi_size_effect$phi, 2)`). The effect size statistic (Phi) suggests, however, a low practical significance for Scar Arrangement and a large pratical significance for Scar Faciality.

@@

plot.morpho.var(morpho_data_opposed_platforms)

>>>>

cohens_area$cohens.f

@@

As in other classes of stone tools, scaled pieces' metric attributes are impacted by both the initial blank size as well as by the intensity of their use. Specifically, because some scaled pieces are used in multiple axes, direct comparisons for length and width of the typological axes cannot be straightforwardly performed. Since the majority of the analyzed pieces presented a rectangular outline, we use the area of a rectangle (i.e. Length x Width) as an approximation for the overall dimensions of the artifacts. Area calculations revealed a maximum of `r round(max(techno_data$Area, na.rm = TRUE), digits = 2)` mm and a minimum of `r round(min(techno_data$Area, na.rm = TRUE), digits = 2)` mm. For thickness the maximum is `r round(max(techno_data$Thickness, na.rm = TRUE), digits = 2)` mm and the minimum is `r round(min(techno_data$Thickness, na.rm = TRUE), digits = 2)` mm. In general, mean values tend to be similar between raw materials, with some differences occurring within the Proto-Solutrean and Magdalenian assemblages, most certainly as a result of the small samples analysed for each of these periods (Figure \@ref(fig:meanplots)). Across techno-complexes, however, significant statistical differences were detected when using an ANOVA test for Thickness (*F* (`r aov_thickness$df[1]`, `r aov_thickness$df[2]`) = `r aov_thickness$statistic[1]`, *p* = `r aov_thickness$p.value[1]`, *d* = `r cohens_thickness$cohens.f`), but not for Area (*F* (`r aov_area$df[1]`, `r aov_area$df[2]`) = `r aov_area$statistic[1]`, *p* = `r aov_area$p.value[1]`, *d* = `r cohens_area$cohens.f`).

@@

In three (Scar Arrangement, Scar Extension and Edge Delineation) of the six represented variables, the category 'Other' is one of the most frequent (above 35%) for both quartz and chert, revealing a very high variability for the combination of attributes within each variable. A chi-square test with modified alpha level (Bonferroni correction) to adjust for multiple testing and reduce type-I error, shows significant differences in the quartz and chert Scar Arrangement categories (*X^2^* [`r chi_arrangement$chi_test$parameter`, *N* = `r sum(chi_arrangement$chi_test$observed)`] = `r  round(chi_arrangement$chi_test$statistic, 2)`, *p* = `r round(chi_arrangement$chi_test$p.value, 3)`, *phi* = `r round(chi_arrangement$chi_size_effect$phi, 2)`), and in Scar Faciality categories (*X^2^* [`r chi_faciality$chi_test$parameter`, *N* = `r sum(chi_faciality$chi_test$observed)`] = `r  round(chi_faciality$chi_test$statistic,2)`, *p* = `r round(chi_faciality$chi_test$p.value, 3)`, *phi* = `r round(chi_faciality$chi_size_effect$phi, 2)`). The effect size statistic (Phi) suggests, however, a medium practical significance for both variables.

@@

morpho.var.plot(morpho_data_opposed_platforms)
